Textmining in Economics

### 2017-10-10 textm.Rmd
typo
update

---
title: "Text as Data"
author: "Franziska Löw"
date: "08. Oktober 2017"
output: ioslides_presentation
---

@@

@@

***
- Most empirical work in economics relies on inherently quantitative data: prices, demand, votes, etc.
- But a large amount of unstructured text is also generated in economic environments: 
    - company reports, 
    - policy committee deliberations, 
    - court decisions, 
    - media articles, 
    - political speeches, etc. 
- One can use such data qualitatively, but increasing interest in treating text quantitatively.

## What is Text?

At an abstract level, text is simply a string of characters.

- Some of these may be from the Latin alphabet ('a', 'A', 'p' and so on) but there may also be: 
    - Punctuation (e.g. !), 
    - Numbers, 
    - Non-alphanumerical characters (e.g. @)

**Key Question:** How can we obtain an informative, quantitative representation of these character strings? This is the goal of text mining.

# Representing Text as Data

## Textual Database

- A single observation in a textual database is called a document $D$.

- The set of documents that make up the dataset is called a corpus.

- We often have covariates associated with each document that are sometimes called metadata.

## Reduce Dimensionality

Textmining analysis take the following steps to reduce dimensionality to a managable size:

1. **Represent Text as Data** Bag-of-words model, n-grams, Text-preprocessing

2. **Statistical Methods**

  - Map $\boldsymbol{C}$ to predict values $\boldsymbol{\hat{V}}$ of unknown outcomes $\boldsymbol{V}$. 
	- E.g. the variable of interest $\boldsymbol{V}$ is an indicator whether the email is spam. The prediction $\boldsymbol{\hat{V}}$ determines whether or not to send the email to a spam filter. 
	- Sometimes the attribute of interest is latent (e.g. topics of a newspaper article)

@@

## Represent text as Data
Convert raw text $D$ to a numerical array $\boldsymbol{C}$:
A basic quantitative representation of a corpus is the bag-of-words model:

@@

## Text Mining and Information

@@

Pre-process strings to obtain a cleaner representation:

1. **Tokenization** splitting of a raw character string into individual elements of interest.

2. **Stopword Removal** The frequency distribution of words in natural languages is highly skewed.

3. **Linguistic Roots** For many applications, the relevant information in tokens is their linguistic root, not their grammatical form. (Stemming or Lemmatizing)

4. **Multi-Word Phrases** Sometimes groups of individual tokens like “donald trump” or “fc bayern” have a specific meaning.

## Example: Twitter

Tweets after Stopword Removal.

```{r, echo=FALSE}
rm(list = ls())
load(file = "~/GitHub/politsentiment_data/out/tweets_18_24.rda")

htmlTable::htmlTable(tweets[1:3,c("text","text_cleaned")], align="l")
```

## Tokenization & Term Frequency

How to reflect the importance of a term? 3 concepts:

@@

***
```{r, include=FALSE}
# load the packages
libs <- c("tidytext","tidyr",
          "plyr","dplyr",
          "htmlTable")
lapply(libs, library, character.only = TRUE)
```

```{r, include=FALSE}
rm(list = ls())
load(file = "~/GitHub/politsentiment_data/out/tweets_18_24.rda")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# get the tweets that are retweeted 
tweets$isretweet <- ifelse(grepl('^RT', tweets$text),TRUE,FALSE)

@@

- high-dimensionality of $c_i$, ($p > n^{train}), requires use of regression techniques appropriate (e.g. $L1$ regularized linear or logistic regression)

>>>>

---
title: "Textmining in Economics"
author: "Franziska Löw"
date: "08. Oktober 2017"
output: ioslides_presentation
---

@@

```{r, include=FALSE}
# load the packages
libs <- c("tidytext","tidyr","readr","lubridate","tm","stm",
          "plyr","dplyr","class","knitr","kableExtra","cldr",'network', 'sna', 'qdap',
          "htmlTable","ggplot2","gridExtra","jsonlite","stringr","scales")
lapply(libs, library, character.only = TRUE)
```

@@

## Introduction

- In economics, empirical work is largely based on quantitative data (e.g. prices, demand, votes, etc.)
- But a large amount of unstructured text is also generated in economic environments (e.g. company reports, court decisions, media articles, etc.)
- At an abstract level, text is simply a string of characters.
    - Latin alphabet ('a', 'A', 'p' and so on)
    - Punctuation (e.g. !), 
    - Numbers, 
    - Non-alphanumerical characters (e.g. @)

**Key Question:** How can we obtain an informative, quantitative representation of these character strings?

## Concepts

- A single observation in a textual database is called a *document*.

- The set of documents that make up the dataset is called a *corpus*.

- Covariates associated with each document are called *metadata*.

In the text analysis ["Bundestagswahl 2017"](https://franziloew.github.io/politsentiment/election.html) we use a corpus of Tweets containing one of the parliamentary parties during the german parliamentary elections 2017:

- 151930 Tweets (without retweets) from 12am to 12pm on the 24-10-2017

- A document is a single Tweet

- Associated Metadata: time, user name, if the user is a news outlet or a "privat" twitter-user, etc.

```{r, include=FALSE}
rm(list = ls())
load(file = "~/GitHub/politsentiment_data/out/tweets_18_24.rda")
# during and after elecction
tweets <- tweets[which(tweets$day == "24"),]
tweets <- with(tweets, tweets[hour(date) >= 12 & hour(date) <= 23, ])
```

```{r, message=FALSE, warning=FALSE, include=FALSE}
# get the tweets that are retweeted 
tweets$isretweet <- ifelse(grepl('^RT', tweets$text),TRUE,FALSE)

***

Text data is inherently high-dimensional, so an analysis of text usually tries to reduce this dimensionality following these steps:

1. Represent the textual database $D$ (**document**) as a numerical array $\boldsymbol{C}$ (**corpus**)
  - *Text-preprocessing, Bag-of-words model*

2. **Statistical Methods** 
  - Map $\boldsymbol{C}$ to predict values $\boldsymbol{\hat{V}}$ of unknown outcomes $\boldsymbol{V}$. 
	- Sometimes the attribute of interest is latent (e.g. topics of a newspaper article)

@@

1. **Tokenization** splitting of a raw character string into individual elements of interest: words, numbers, punctuation.

2. **Stopword Removal** Stopwords (highly frequent terms) are stripped out of token lists as they do not help distinguish one document from another.

3. **Linguistic Roots** For many applications, the relevant information in tokens is their linguistic root, not their grammatical form (e.g. treat "looking", "look", "looks" as equivalent tokens)

4. **Multi-Word Phrases** Sometimes groups of individual tokens like “United Kingdom”  have a specific meaning (Use Ngrams)

## Cleaning Data

```{r, echo=FALSE}
htmlTable(orig[1:4,c("text","text_cleaned")], align = "l")
```

## Tokenization

```{r, echo=FALSE, message=FALSE, warning=FALSE}
orig$text_cleaned <- removeWords(orig$text_cleaned, c("cdu","csu","union","spd","fdp","grüne","grünen","linke","afd"))

orig %>%
  #group_by(partei) %>%
  unnest_tokens(word, text_cleaned) %>%
  count(word, sort = TRUE)  %>%
  top_n(15) %>%
  htmlTable()
```

## Tokenization (Bigrams)

```{r, echo=FALSE, message=FALSE, warning=FALSE}
orig %>%
  #group_by(partei) %>%
  unnest_tokens(bigram, text_cleaned, token="ngrams", n=2) %>%
  count(bigram, sort = TRUE)  %>%
  top_n(15) %>%
  htmlTable()
```

## What is the data?

After pre-processing, each document is a finite list of terms.

A basic quantitative representation of a corpus is the **bag-of-words model**

@@

## bag-of-words model

@@

# Term Frequency

***
How to reflect the importance of a term? 3 concepts:

@@

***

@@

- high-dimensionality of $c_i$, ($p > n^{train}$), requires use of regression techniques appropriate (e.g. $L1$ regularized linear or logistic regression)
